\documentclass[a4paper,twoside]{article}   
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphics}
\usepackage{epsfig}
% 
%  
% Please submit at http://www.easychair.org/conferences/?conf=ttc2011. Your submission should include (i) a case description (in PDF format) answering the above questions, (ii) a ZIP archive that contains all test artifacts, and (iii) an evaluation scheme (a spreadsheet file or a pointer to an online "classification scheme").

\begin{document}

\title{Submission for Transformation Tool Contest 2011: Sample URDAD model}

\author{ Fritz Solms, Craig Edwards and Stefan Gruner\\
  {\small Dept. of Comp. Science, University of Pretoria, email: fritz@solms.co.za, sgruner@cs.up.ac.za}}

\maketitle

\abstract{The Use-Case, Responsibility-Driven Analysis and Design (URDAD) method, is a method used by requirements specialists (typically business analysts) to perform technology neutral, services oriented analysis and business process design. The output of the URDAD method is an URDAD model containing the requirements and technology neutral process design which was historically encoded in UML. In order to improve the productivity and quality of technology neutral analysis and design a formal metamodel and a domain specific language with a concrete text syntax have been developed. The submission provides the URDAD metamodel as well as an example model for a typical business use case. The modl is technology neutral and does not prescribe the implementation technologies and architecture. The transformation should preferably target a typical enterprise systems architecture like the Java Enterprise Edition (JavaEE) reference architecture or a Services Oriented Architecture.}

%\include{abstract}

\tableofcontents

\include{contextualization}

\include{providedArtifacts}

\include{useCaseDescription}

\include{useCaseVariationPoints}

\include{implementationMappingGuidelines}

\include{evaluationGuidelines}

\bibliographystyle{plain}  %%abbrv
\bibliography{../../bibliography}

\end{document}

