\section{Related work \label{sec:relatedWork}}

The URDAD methodology provides a service-oriented methodology for generating a semi-formal analysis and design model representing MDA's PIM and supporting test and implementation generation. The work reported in \cite{iacob_model-driven_2008} discusses an alternative approach. Business Rules are specified using OMG's {\em Semantics for Business Vocabulary and Rules} (SBVR). Using MDA tools, these are mapped onto service specifications and orchestration and BPEL process specifications. Processes are thus assembled from services which are related to business rules. This is similar to our satisfiability links specifying the services used to realize the different functional requirements.

The URDAD DSL allows for the specification of textual and graphical grammars through which the URDAD model is populated. An alternative approach is to define a separate metamodel for the use case narrative and to transform the narrative use case model onto a UML based requirements model \cite{hoffmann_towards_2009,osis_transforming_2010}. This approach introduces the complexities of having two languages with potential inconsistency risks and an additional model transformation. It thus requires extensive consistency checks between the narrative and the UML models.

\cite{asnina_computation_2010} stress the need of modelling in the problem domain as well the benefits of accumulating requirements within a single model. Services are grouped into feature sets which are related to responsibility domains. Functional requirements are decomposed across levels of granularity and higher level processes are orchestrated across lower level services. They define the notion of functionals with cause and effect which can be related to the concept of a service contract. In addition they provide a {\em topological functional model} (TFM) for mapping technology neutral service requirements onto the available concrete service pool. The TFM and in particular the mapping of service requirements onto concrete services is independent of the modelling technique and can be applied to an URDAD model. 

In the {\em Requirements Driven Design Automation} methodology (RDDA) \cite{cardei_model_2008} one encodes requirements specifications in SYSML diagrams. The SYSML model is enriched with semantic descriptions after which the model is transformed to the {\em One Pass to Production} (OPP) design language, the ODL. ODL is an OWL based ontology from which the requirements are validated for consistency and completeness. The approach is, however, structure focused with little emphasis on service contracts and recursive orchestration of higher level services from lower level services.
