\abstract{
URDAD is a methodology which is used by requirements engineering to develop semi-formal requirements models. Historically these requirements models were encoded in UML. However, capturing the requirements model coming out of the URDAD process in UML requires a high level of skill and discipline and includes tedious tasks around inserting certain semantic relationships required by the URDAD method. The resultant process is inefficient and error prone. 

The resultant process is inefficient and error prone. This paper introduces a domain-specific language supporting the URDAD methodology. We introduce and assess an URDAD metamodel which supports URDAD model encodings with significantly lower language and model+complexity than an URDAD UML profile. Also, model validation for completeness and consistency as well as

  	12 	

+URDAD tool development are simplified. We also present simple concrete

  	13 	

+text and diagrammatic syntaxes for URDAD, which capture the URDAD semantics in a normalized way. These syntaxes can automatically be derived from the URDAD meta-model. 

To enforce model quality for a UML encoding of an URDAD model one needs to specify an URDAD UML profile introducing certain URDAD-specific semantics and extensive OCL constraints which constrain the UML model to a valid, consistent and ultimately complete URDAD model. This paper presents and assesses the alternative approach of introducing a domain-specific language (DSL) supporting the URDAD method and defines a concrete text syntax which can be used to encode an URDAD requirements model. The URDAD-DSL provides simpler language to encode an URDAD model, results in models with lower model complexity, enforces the URDAD model structure, facilitates faster model encoding and results in higher quality models. Model validation for completeness and consistency and URDAD tool development are thus simplified.
}
 

