\section{Abstract}

\begin{frame}{Problem Specification}
\initclock

  \begin{itemize}
    \item<+-| alert@+> Inferior requirements
      \begin{itemize}
      	\item Core contributor to poor software quality \& high cost.
      \end{itemize}
    \item<+-| alert@+> Formal methods
      \begin{itemize}
	      \item Use mathematical modeling \& formal logic to to specify \& verify requirements.
    	  \item However, incur high cost \& skills requirements.
      \end{itemize}
    \item<+-| alert@+> Semi-formal methods
      \begin{itemize}
		\item Constrain cost \& skills requirements.
		\item Formalization of process \& inputs/outputs.
	  		\note[item]{Main complexity in constraint specification.}
	  \end{itemize}		
   \item<+-| alert@+> Model Driven Engineering (MDE)
   	  \begin{itemize}
   	  	\item fall into class of semi-formal approaches
		\item Often no defined, repeatable engineering process.
		\item Design structures, quality \& semantics often vary.
	  \end{itemize}	
  \end{itemize}

\end{frame}

%----------------------------------------------------%

\begin{frame}{Novel Contribution}

  \begin{itemize}
   \item<+-| alert@+> Understanding of quality crite
   \item<+-| alert@+> 
   \item<+-| alert@+> 
   \item<+-| alert@+> 
   \item<+-| alert@+> 
  \end{itemize}

\end{frame}

%----------------------------------------------------%

\begin{frame}{Abstract}
\initclock

  \begin{itemize}
   \item<+-| alert@+> URDAD is a semi-formal, service-oriented A\&D methodology 
      \begin{itemize}
	\item generates technology neutral requirements model (PIM),
	\item supported by metamodel \& DSL
      \end{itemize}
    \item<+-| alert@+> We identify
    \begin{itemize}
    \item stakeholders quality requirements 
	\begin{itemize}
	  \item for process \& resulting model,
	\end{itemize}
    \item quality drivers used to address each quality requirement,
    \end{itemize}
      \item<+-| alert@+> and show
      \begin{itemize}
	\item how quality drivers are embedded in URDAD,
	\item internal consistency of URDAD.
      \end{itemize}
  \end{itemize}

\end{frame}
